
---
title: "Experimental Design"
author: "Jeremy"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
- Key Components of an Experiment.
1. Randomization
2. Replication
3. Blocking

Randomization - evenly distributes any variability in outcome due to outside factors across treatments groups.

Replication - must repeat an experiment to fully assess variability. e.g. If we conduct drug efficacy on
one person, how can we properly generalize those results? (We can't)

Blocking - helps control variability by making treatment groups more alike. Inside of groups, differences will be minimal, and across groups, differences will be large.

```{r cars}
library(broom)
library(tidyverse)
library(pwr)
library(readr)
library(stats)
library(recipes)
```

# Perform a t-test

```{r pressure, echo=FALSE}
ToothGrowth_ttest <- t.test(len ~ supp, data = ToothGrowth)
summary(ToothGrowth_ttest)

tidy(ToothGrowth_ttest)
```

# Count number of observations for each combination of supp and dose

```{r}
ToothGrowth %>% 
  count(supp, dose)
```

# Create a boxplot with geom_boxplot()

```{r}
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) +
  xlab("supp")
```

# One way ANOVA table

```{r}
ToothGrowth_aov <- aov(len ~ dose + supp, data = ToothGrowth)

# Examine ToothGrowth_aov
summary(ToothGrowth_aov)
```

# Hypothesis testing
NULL hypothesis - There is no change, no differences between groups, and mean, nedian, or observation = a number.

ALT hypothesis - There is a change, difference in groups, and mean, median, or observation is >,<,!= to a number

POWER - Is the probability that the test correctly rejects null hypothesis when the alternative hypothesis is true.

EFFECT SIZE - standardized measure of the difference you're trying to detect.

SAMPLE SIZE - How many experimental units you need to survey to detect the desired difference at the desired power.

# Interpretation
* There is a change or difference in groups since the mean is less than 18.

```{r}
# Less than
t.test(ToothGrowth$len, alternative = "less", mu = 18)
```

# Interpretation
* The mean is greater than 18 thus there is a change or difference between groups.

```{r}
# Greater than
t.test(ToothGrowth$len, alternative = "greater", mu = 18)
```

# Test statistics using pwr library
- One argument has to be NULL for it to be calculated by the call to pwr.*() functions.
- A call to any pwr.*() function returns an object of class power.htest, which can then manipulated in the same way as R objects.
- One key part of designing an experiment is knowing the required sample size you'll need to be able to test your hypothesis.

In the code chunk below, the output power is 0.2521043, therefore it is not great, we probably can't detect that small of an effect size with so few people in each group.


```{r}
pwr.anova.test(k = 3,
               n = 20,
               f = 0.2,
               sig.level = 0.05,
               power = NULL)
```


```{r}
# Calculate power
pwr.t.test(n = 100, 
           d = 0.35,
           sig.level = 0.10,
           type = "two.sample", 
           alternative = "two.sided",
           power = NULL)
```

```{r}
# Calculate sample size
pwr.t.test(n = NULL, 
           d = 0.25, 
           sig.level = 0.05, 
           type = "one.sample", alternative = "greater", 
           power = 0.8)
```

```{r, echo=FALSE}
lendingclub <- read_csv("C:/Users/jerem/Downloads/archive/loan.csv")
# names(lendingclub)
```


```{r}
# Find median loan_amnt, mean int_rate, and mean annual_inc
lendingclub %>% summarize(median(loan_amnt), mean(as.numeric(int_rate)), mean(annual_inc))
```

S
```{r}
# Use ggplot2 to build a bar chart of purpose
ggplot(data=lendingclub, aes(x = purpose)) + 
  geom_bar() +
  coord_flip()
```


```{r}
# Use recode() to create the new purpose_recode variable
lendingclub$purpose_recode <- lendingclub$purpose %>% recode( 
  "credit_card" = "debt_related", 
  "debt_consolidation" = "debt_related",
  "medical" = "debt_related",
  "car" = "big_purchase", 
  "major_purchase" = "big_purchase", 
  "vacation" = "big_purchase",
  "moving" = "life_change", 
  "small_business" = "life_change", 
  "wedding" = "life_change",
  "house" = "home_related", 
  "home_improvement" = "home_related")


# Build a linear regression model, purpose_recode_model
purpose_recode_model <- lm(funded_amnt ~ purpose_recode, data = lendingclub)

# Examine results of purpose_recode_model
summary(purpose_recode_model)
```

# ANOVA
- Used to compare 3+ groups.
- Can be implemented in 2 ways using anova() and aov().

```{r}
# Get anova results and save as purpose_recode_anova
purpose_recode_anova <- anova(purpose_recode_model)

# Print purpose_recode_anova
print(purpose_recode_anova)

# Examine class of purpose_recode_anova
class(purpose_recode_anova)
```

- The second way of implementing anova:

```{r}
# Use aov() to build purpose_aov
purpose_aov <- aov(funded_amnt ~ purpose_recode, data = lendingclub)

# Conduct Tukey's HSD test to create tukey_output
tukey_output <- TukeyHSD(purpose_aov, "purpose_recode", conf.level = 0.95)

# Tidy tukey_output to make sense of the results
tidy(tukey_output)
```


```{r}
# Use aov() to build purpose_emp_aov
purpose_emp_aov <- aov(funded_amnt ~ purpose_recode + emp_length, data = lendingclub)

# Print purpose_emp_aov to the console
purpose_emp_aov

# Call summary() to see the p-values
summary(purpose_emp_aov)
```


```{r}
# Convert percentage strings to numeric values
lendingclub$int_rate <- as.numeric(gsub("%", "", lendingclub$int_rate)) 

lendingclub$grade <- factor(lendingclub$grade)

# Examine the summary of int_rate
summary(lendingclub$int_rate)

# Examine int_rate by grade
lendingclub %>% 
  group_by(grade) %>%
  summarize(mean = mean(int_rate), var = var(int_rate), median = median(int_rate))

# Make a boxplot of int_rate by grade
ggplot(lendingclub, aes(x = grade, y = int_rate)) + 
  geom_boxplot()
```


```{r}
lendingclub_blu <- lendingclub %>% recipe(int_rate ~ grade) %>% step_impute_mode()

lendingclub_cleaned <- lendingclub_blu %>% prep(lendingclub) %>% bake(lendingclub)
```

# Plot of grade_aov interpretation
- The "Residuals vs Fitted Plot" will show if the model is a good fit, if there is a similar scatter pattern for each level of the group variable. If there was a different pattern, we could begin to think there's heteroscedasticity in the residuals, and the model might not be a good fit.
- The "Normal Q-Q Plot" should, ideally, show the points scattered around the regression line. One assumption of ANOVA and linear models is that the residuals are normally distributed for the model to be a good fit.
- A good fit will show in your "Scale-Location Plot" as the residuals increasing with the fitted values, we see that here.
- The "Residuals vs Leverage Plot" shows which levels are the best fitted to the model. Here the smaller levels seem better fit.

```{r}
# Fit the ANOVA model
grade_aov <- aov(int_rate ~ grade, data = lendingclub_cleaned)

# Display ANOVA summary
summary(grade_aov)

# For a 2x2 grid of plots:
par(mfrow=c(2, 2))

# Plot grade_aov
plot(grade_aov)

# Bartlett's test for homogeneity of variance
bartlett.test(int_rate ~ grade, data = lendingclub_cleaned)

# Conduct the Kruskal-Wallis rank sum test
kruskal.test(int_rate ~ grade,
             data = lendingclub_cleaned)
```

- We need to check if the variances across groups are different.
- When different proceed with assumption of anova and linear modeling i.e. homogeneity of variances using barttlett.test(), which takes a formula and a dataset.
- Another alternative to anova is Kruskal-Wallis rank sum test. It is an extension of the Mann-Whitney U test for when there are more than two groups, like the "grade" variable. For us, the null hypothesis for this test would be that all of the "init_rate" have the same ranking by "grade".

```{r}
# Load the pwr package
library(pwr)

# Use the correct function from pwr to find the sample size
pwr.t.test(d = 0.2, 
           n = NULL, 
           sig.level = 0.05, 
           alternative = "two.sided", 
           power = 0.8)
```


```{r}
lendingclub_ab <- lendingclub %>% filter(grade == c("A","B"))

# Plot the A/B test results
ggplot(lendingclub_ab, aes(x = grade, y = loan_amnt)) + 
  geom_boxplot()

# Conduct a two-sided t-test
t.test(loan_amnt ~ grade, data = lendingclub_ab)

# Build lendingclub_multi
lendingclub_multi <- lm(loan_amnt ~ grade + verification_status, lendingclub_ab)

# Examine lendingclub_multi results
tidy(lendingclub_multi)
```

